
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.
```{r}
library(readr)
library(tidyverse)
df <-read_csv("C:/Users/student/Desktop/421math/math241assignments/adult_census_missing_right.csv")
colSums(is.na(df))

```


2. Using function `aggr` the VIM package to plot the number of NAs for each column.
```{r, eval=FALSE}
library(tidyverse)
library(VIM)

aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
colSums(df=='Unknown', na.rm = TRUE)
table(df$education)
table(df$age)
table(df$workclass)
table(df$marital.status)
table(df$occupation)
table(df$sex)
table(df$capital.gain)
table(df$native.country)
```


4. Replace all the forms of missing values found to NA
```{r}
library(tidyverse)
df <- replace(df, 
              df == 'Unknown' |
              df == '?' |
              df == 'Not Known', NA)
```



5. Replot the number of NAs for each column. 
```{r}
barplot(colSums(is.na(df)))
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
df1 = drop_na(df)
```


7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
df2 <- df %>% 
  fill(age, workclass, education, marital.status, occupation, sex, capital.gain, native.country, .direction = "updown")
```



8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
df3 <- df
#Numeric Variables
df3$age <- replace_na(df$age,29)

df3$capital.gain <- replace_na(df3$capital.gain, median(df3$capital.gain, na.rm=TRUE))


#Categorical Variables with majority values 

majority_workclass <- names(which.max(table(df3$workclass)))
df3$workclass <- replace_na(df3$workclass, majority_workclass)

majority_education <- names(which.max(table(df3$education)))
df3$education <- replace_na(df3$education, majority_education)

majority_marital_status <-  names(which.max(table(df3$marital.status)))
df3$marital.status <- replace_na(df3$marital.status, majority_marital_status)

majority_occupation <- names(which.max(table(df3$occupation)))
df3$occupation <- replace_na(df3$occupation, majority_occupation)

majority_sex <- names(which.max(table(df3$sex)))
df3$sex <- replace_na(df3$sex, majority_sex)

majority_native_country <- names(which.max(table(df3$native.country)))
df3$native.country <- replace_na(df3$native.country, majority_native_country)

```






