
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 

1. List all the names of the columns of the data
```{r}
library(dplyr)
df = read.csv("C:/Users/student/Desktop/421math/math241assignments/imdb_top_1000.csv")
names(df)
```


2. Which movies have the highest money earned (Gross)?
```{r}
df %>%
  arrange(desc(Gross)) %>%
  select(Series_Title, Gross)
```

3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.
```{r}

lowest_rating <- min(df$IMDB_Rating)
df %>%
filter(IMDB_Rating == lowest_rating) %>%
select(Series_Title, IMDB_Rating) %>%
head(5)
```


4. Which year have the most number of movies released in the list?  What is the total of money earned on that year?
```{r}
df %>%
count(Released_Year, sort = TRUE)
```
```{r}
df %>% 
  mutate(Gross = as.numeric(gsub(",", "", Gross))) %>%
  group_by(Released_Year) %>%
  summarise( 
    num_movies = n(), 
    total_gross = sum(Gross, na.rm = TRUE) 
  ) %>%
  arrange(desc(num_movies)) %>%
  slice(1)
```


5. What is the total money earned per movies on average?
```{r}
df %>%
  mutate(Gross = as.numeric(gsub(",", "", Gross))) %>%
  summarise(avg_gross = mean(Gross, na.rm = TRUE))
```


6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.
```{r}
df %>%
  group_by(Released_Year) %>% 
  summarise(avg_votes = mean(No_of_Votes, na.rm = TRUE)) %>%
  arrange(desc(avg_votes))
```
```{r}
df %>% 
  filter(IMDB_Rating > 9) %>%
  summarise(avg_votes = mean(No_of_Votes, na.rm = TRUE))
```


7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile. 
```{r}
df %>%
filter(
Released_Year == 2020,
No_of_Votes >= quantile(No_of_Votes, 0.50, na.rm = TRUE),
No_of_Votes < quantile(No_of_Votes, 0.75, na.rm = TRUE)
) %>%
summarise(avg_meta_score = mean(as.numeric(Meta_score), na.rm = TRUE))
```


8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=FALSE}
# Select only character columns
df %>% select_if(is.character)

# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```
```{r}
df %>% 
  summarise_if(is.numeric, mean, na.rm =TRUE)
df %>% 
  summarise_if(is.numeric, max, na.rm = TRUE)
```


10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
```{r}
df %>%
  select(Series_Title, Released_Year, IMDB_Rating)
#Displays only the movie title, released year, IMDB rating
```

- filter
```{r}
df %>% 
  filter(IMDB_Rating >= 9)
#Shows movies with an IMDb rating of 9 or higher 
```

- mutate
```{r}
df %>% 
  mutate(Rating_Category = 
           ifelse(IMDB_Rating >= 8.5, "Excellent", "Good"))
#Creates a new column that categorizes moves based on IMDb rating
```

- summarise
```{r}
df %>% 
  summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE))
#calculates the average IMDb rating over all of the movies
```

- arrange
```{r}
df %>% 
  arrange(desc(IMDB_Rating))
#Sorts movies from highest to lowest by IDMb Rating
```

- count
```{r}
df %>% 
  count(Certificate)
#counts the number of movies belonging to each certificate category
```

- count + arrange
```{r}
df %>% 
  count(Genre) %>% 
  arrange(desc(n))
#shows the most common genres in the dataset 
```

- filter + count + arrange
```{r}
df %>% 
  filter(IMDB_Rating >= 8.5) %>%
  count(Released_Year) %>% 
  arrange(desc(n))
#identifies which years produced the most highly rated movies over 8.5
```

- group_by + summarise
```{r}
df %>%
group_by(Released_Year) %>%
summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE))
#calculates the average IMDb Rating for each relase year
```

- filter + group_by + summarise
```{r}
df %>% 
  filter(Released_Year >= 2010) %>% 
  group_by(Released_Year) %>% 
  summarise(avg_votes = mean(No_of_Votes, na.rm = TRUE))
# calculates the average number of votes for movies released in 2010 or later 
```

- filter + group_by + summarise + arrange
```{r}
df %>%
  filter(Released_Year >= 2010) %>%
  group_by(Released_Year) %>%
  summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE)) %>%
  arrange(desc(avg_rating))
#Ranks years since 2010 by thier average IMDb rating, from highest to lowest 
```

