---
title: "Assignment 3"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment3.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28, 34, 29, 42), 
                Name = c('Tom', 'Jack', 'Steve', 'Ricky'))

```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://bryantstats.github.io/math421/data/us-states.csv).

    ```{r}
    df = read.csv("C:/Users/student/Desktop/421math/math241assignments/us-states (1).csv")
    ```

3.  Show the names of the variables in the data

    ```{r}
    names(df)

    ```

```         
```

4.  How many columns and rows the data have?

    ```{r}
    dim(df)
    ```

5.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?

    ```{r}
    sum(is.na(df))
    colSums(is.na(df))
    ```

There are no missing values in this dataset.

6.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.

    ```{r}
    class(df$date)
    df$date = as.Date(df$date)
    class(df$date)

    ```

    It is originally showing as "character" after the code, it is shown as date

7.  Capitalize the names of all the variables

    ```{r}
    names(df) = toupper(names(df))
    names(df)

    ```

    ```{r}
    names(df) = tolower(names(df))
    ```

8.  Find the average number of cases per day. Find the maximum cases a day.

    ```{r}
    mean(df$cases)
    max(df$cases)
    ```

9.  How many states are there in the data?

    ```{r}
    length(unique(df$state))

    ```

10. Create a new variable `weekdays` to store the weekday for each rows.

    ```{r}
    df$date = as.Date(df$date)
    df$weekdays = weekdays(df$date)
    head(df)

    ```

11. Create the categorical variable `death2` variable taking the values as follows

-   `has_death` if there is a death that day

-   `no_death` if there is no death that day

    ```{r}
    df$death2 = ifelse(df$deaths > 0, "has_death", "no_death")
    head(df)

    ```

Find the frequency and relative frequency of `no_death` and `has_death`.

```{r}
table(df$death2)

prop.table(table(df$death2))

```

12. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

    ```{r}
    summary(df$deaths)

    ```

    Q1: 838 Q2: 4064 Q3: 12674

13. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3

    ```{r}
    df$death3 = ifelse(df$deaths < 838, "low_death", 
                       ifelse(df$deaths > 12674, "high_death", "mid_death"))
    table(df$death3)
    ```

14. Find the average cases in Rhode Island in 2021

    ```{r}
    mean(df$cases[df$state == "Rhode Island" & df$date >= "2021-01-01" & df$date <= "2021-12-31"])
    ```

15. Find the median cases by weekdays in Rhode Island in 2021

    ```{r}
    ri = df[df$state == "Rhode Island" & format(df$date, "%Y") == "2021", ]

    tapply(ri$cases, ri$weekdays, median)
    ```

16. Compare the median cases in Rhode Island in June, July, August and September in 2021.

    ```{r}
    ri = df[df$state == "Rhode Island" & df$date >= "2021-06-01" & df$date <= "2021-09-30", ]

    tapply(ri$cases, format(ri$date, "%B"), median)
    ```

17. Find your own dataset, import it and implement the following functions on the data

    ```{r}
    titanic = read.csv("C:/Users/student/Downloads/titanic.csv")
    library(dplyr)
    ```

-   head

    ```{r}
    head(titanic)

    ```

-   str

    ```{r}
    str(titanic)
    ```

-   names

    ```{r}
    names(titanic)
    ```

-   mean, min, max

    ```{r}
    mean(titanic$Age, na.rm = TRUE)
    min(titanic$Age, na.rm = TRUE)
    max(titanic$Age, na.rm = TRUE)
    ```

-   table

    ```{r}
    table(titanic$Survived)
    ```

-   is.na

    ```{r}
    sum(is.na(titanic))
    ```

-   colSums

    ```{r}
    colSums(is.na(titanic))
    ```

-   class

    ```{r}
    class(titanic$Age)
    ```

-   cor

    ```{r}
    cor(titanic$Age, titanic$Fare, use = "complete.obs")

    ```

-   by

    ```{r}
    by(titanic$Fare, titanic$Sex, mean, na.rm = TRUE)
    ```

-   ifelse

    ```{r}
    titanic$SurvivalStatus = ifelse(titanic$Survived == 1, "Survived", "Did not survive")
    table(titanic$SurvivalStatus)
    ```

-   case_when

    ```{r}

    titanic$AgeGroup = case_when(
      titanic$Age < 18 ~ "Child",
      titanic$Age < 60 ~ "Adult",
    TRUE ~ "Senior"
    )
    table(titanic$AgeGroup)
    ```

18. In the dataset in #17, practice the follows. You can reuse the code of 17.

-   Create a categorical variable from a continuous variable

    ```{r}
    library(dplyr)
    titanic$AgeGroup = case_when(
      titanic$Age < 18 ~ "Child",
      titanic$Age < 60 ~ "Adult",
      TRUE ~ "Senior"
    )
    table(titanic$AgeGroup)
    ```

-   From one categorical variable, create a new categorical variable with fewer categories

    ```{r}
    titanic$ClassGroup = ifelse(titanic$Pclass == 1, "Higher Class", "Lower Class"
    )
    table(titanic$ClassGroup)
    ```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](https://bryantstats.github.io/math421/data/titanic.csv)
