
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.
```{r}
library(readr)
library(tidyverse)
df <-read_csv("C:/Users/student/Desktop/421math/math241assignments/adult_census_missing_right.csv")
colSums(is.na(df))

```


2. Using function `aggr` the VIM package to plot the number of NAs for each column.
```{r, eval=FALSE}
library(tidyverse)
library(VIM)

aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 
```{r}
table(df$education)
sum(is.na(df$age))
table(df$workclass)
table(df$marital.status)
table(df$occupation)
table(df$sex)
table(df$capital.gain, useNA = "always")
table(df$native.country)
```


4. Replace all the forms of missing values found to NA
```{r}
library(tidyverse)
df <- replace(df, df == 'Unknown' |
                  df == '?'       |
                  df == 'Not Known', NA)
```
```{r}


```


5. Replot the number of NAs for each column. 

6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?
```{r}
df1 = drop_na(df)
```


7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 
```{r}
df2 <- df %>% 
  fill(age, workclass, education, marital.status, occupation, sex, capital.gain, native.country, .direction = "updown")
```



8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
mean_age <- mean(df$age, na.rm = TRUE)
df$age <- replace_na(df$age, mean_age)

mean_fnlwgt <- mean(df$fnlwgt, na.rm = TRUE)
df$fnlwgt <- replace_na(df$fnlwgt, mean_fnlwgt)

mean_education_num <- mean(df$education.num, na.rm = TRUE)
df$education.num <- replace_na(df$education.num, mean_education_num)

mean_capital_gain <- mean(df$capital.gain, na.rm = TRUE)
df$capital.gain <- replace_na(df$capital.gain, mean_capital_gain)

mean_capital_loss <- mean(df$capital.loss, na.rm = TRUE)
df$capital.loss <- replace_na(df$capital.loss, mean_capital_loss)

mean_hours <- mean(df$hours.per.week, na.rm = TRUE)
df$hours.per.week <- replace_na(df$hours.per.week, mean_hours)
```






