---
title: "Assignment 4"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment4q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(lubridate)
```

------------------------------------------------------------------------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so.

```{r, eval=FALSE}
install.packages('tidyverse')
```

------------------------------------------------------------------------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://bryantstats.github.io/math421/data/all-states-history.csv). Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv.

```{r}
df = read_csv("C:/Users/student/Downloads/all-states-history.csv")
```

------------------------------------------------------------------------

#### 3. Fix the date and ceate some new variables

`lubridate` is a package of the `tidyverse` packages. We will make uses of `lubridate` in this question.

-   Use the below codes to create `month`, `weekday` and `monthday` variables

```{r, eval=TRUE}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

------------------------------------------------------------------------

#### 4. Create new variables with `case_when`.

The function `case_when` is a good option to create a new variable from existing variable. For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable. `deathIncrease` is the number of daily new death by Covid19. The new variable `daily_death` takes three values: low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14). Please notice that this can also be done in a different way as shown in Assignment 3.

```{r, eval=TRUE}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
)
```

-   Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month \> 20).

    ```{r}
    nrow(df)
    length(df$date)
    month(df$date)[1:10]
    df$month2 <- case_when(
      df$monthday <= 10 ~ "earlymonth",
      df$monthday <= 20 ~ "mid_month", 
      TRUE ~ "end_month"
      )
    ```

-   Create variable `weekend` that takes two values: 1 if it's Saturday or Sunday and 0 otherwise.

    ```{r}
    library(dplyr)
    df$weekend <- case_when(
      df$weekday %in% c(1,7) ~ 1, 
      TRUE ~ 0
    )
    ```

------------------------------------------------------------------------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.

```{r}
df <- df %>% 
  select(-totalTestsViral)
```

------------------------------------------------------------------------

#### 6. Pipe Operator ( %\>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable. Pipe works very well with all the `tidyverse` packages. Refer to [these slides](https://bryantstats.github.io/math421/slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
sqrt(x)

sum(sqrt(x))

log(sum(sqrt(x)))

# log base 2 of 16
log(16, 2)

x <- c(1:10)

x %>% 
  sqrt() %>%
  sum() %>%
  log()

```

------------------------------------------------------------------------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

```{r, eval=TRUE}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))

```

-   Calculate the median number of cases (`positiveIncrease`) by `month`

    ```{r}
    df %>%
    group_by(month) %>%
    summarise(median(positiveIncrease))
    ```

-   Calculate the average number of cases (`positiveIncrease`) by `month2`

    ```{r}
    df %>%
    group_by(month2) %>%
    summarise(mean(positiveIncrease))
    ```

-   Calculate the median number of cases (`positiveIncrease`) by `weekend`

    ```{r}
    df %>%
    group_by(weekend) %>%
    summarise(median(positiveIncrease))
    ```

------------------------------------------------------------------------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r, eval=TRUE}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020.

    ```{r}
    df %>%
    filter(weekend == 1, month %in% c(10, 11)) %>%
    group_by(month) %>%
    summarise(median_cases = median(positiveIncrease, na.rm = TRUE))
    ```

-   Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020

    ```{r}
    df %>%
      filter(month %in% c(9, 10, 11)) %>%
      group_by(month2) %>%
      summarize(avg_deaths = mean(deathIncrease, na.rm = TRUE))
    ```

-   Compare the average number of hospitalizations between weekdays and weekends in Summer 2020

    ```{r}
    df %>% 
      filter(month %in% c(6, 7, 8)) %>%
      group_by(weekend) %>%
      summarise(avg_hosp = mean(hospitalizedCurrently, na.rm = TRUE))
    ```

-   Redo Questions 14 and 15 in [Assignment 3](assignment3q.html) using the combos. Notice: you also need to use the data used in Assignment 3.

    ```{r}
    library(tidyverse)
    old <- read.csv("C:/Users/student/Desktop/421math/math241assignments/us-states (1).csv")
    #14
    old$weekdays <- weekdays (as.Date(old$date))
    old %>%
    filter(state == "Rhode Island",
          date >= "2021-01-01",
          date <= "2021-12-31") %>%
    summarise(avg_cases = mean(cases, na.rm = TRUE))
    #15
    old %>%
      filter(state == "Rhode Island",
            format(as.Date(date), "%Y") == "2021") %>%
    group_by(weekdays) %>%
    summarise(median_cases = median(cases, na.rm = TRUE))
    ```

------------------------------------------------------------------------

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.

```{r}
df %>%
  filter(weekend == 1) %>%
  group_by(month) %>%
summarise(total_deaths = sum(deathIncrease, na.rm = TRUE)) %>%
  arrange(desc(total_deaths))
```

------------------------------------------------------------------------

#### 10. Use your own dataset and implement the follows functions or combos.

You can use the [Adult Census Income](https://bryantstats.github.io/math421/data/adult_census.csv) or [Titanic data](https://bryantstats.github.io/math421/data/titanic.csv).

-   select

    ```{r}
    aci <- read_csv("C:/Users/student/Downloads/adult_census.csv")
    aci %>%
      select(age, education, income)

    ```

-   filter

    ```{r}
    aci %>%
      filter (age > 50)
    ```

-   mutate

    ```{r}
    aci %>%
      mutate(age_group = ifelse(age >= 50, "Older", "Younger"))
    ```

-   summarise

    ```{r}
    aci %>%
      summarise((avg_age = mean(age)))
    ```

-   arrange

    ```{r}
    aci %>%
      arrange(desc(age))
    ```

-   count

    ```{r}
    aci %>% 
      count(income)
    ```

-   count + arrange

    ```{r}
    aci %>% 
      count(education) %>%
      arrange(desc(n))
    ```

-   filter + count + arrange

    ```{r}
    aci %>% 
      filter(age >50) %>% 
      count(occupation) %>% 
      arrange (desc(n))
    ```

-   group_by + summarise

    ```{r}
    aci %>%
      group_by(income) %>%
      summarise(avg_age = mean(age))
    ```

-   filter + group_by + summarise

    ```{r}
    aci %>% 
      filter (sex == "Female") %>% 
      group_by(income) %>% 
      summarise(avg_age = mean(age))
    ```

-   filter + group_by + summarise + arrange

    ```{r}
    aci %>% 
      filter(income == ">50K") %>%
      group_by(occupation) %>% 
      summarise(avg_age = mean(age)) %>% 
      arrange(desc(avg_age))
    ```
